### 2022-06-13 paper.tex
fmt abstract
moved figrues

\usepackage[a4paper,hmargin=1.5cm,vmargin=2.5cm,]{geometry}
\setlength{\columnsep}{0.7cm}
%\usepackage{microtype} % nice typesetting  % doesn't work for me :-( -S
\usepackage{amsmath}
\usepackage{graphicx}

\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\urlstyle{same} % no tt font for URLs
\usepackage{natbib}
\bibliographystyle{genome_research}
%\setcitestyle{aysep={}} 
\usepackage[dvipsnames]{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={blue!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!50!black}
}
\renewcommand{\dbltopfraction}{0.9}
\renewcommand{\textfraction}{0.15}

@@

June 2022
\vspace{6ex}
}]
\blfootnote{\hspace{-.5cm}\raggedright\textsuperscript{*} e-mail:  \texttt{l.kremer@dkfz-heidelberg.de} and\\ \hspace{1cm}\texttt{simon.anders@bioquant.uni-heidelberg.de}}

\section{Abstract}

Single-cell-bisulfite sequencing (scBS) allows to asses DNA methylation with single-basepair and single-cell resolution. Analysing the big data sets thus obtained requires preprocessing to reduce data size, improve signal over noise, and provide interpretability. Currently, this is simply done by tiling the genome into large windows and averaging in these. We show that this coarse-graining dilutes signal, and suggest improved approaches to determine informative regions to quantify for and to quantitate methylation in these windows. We demonstrate how our preprocessing approach helps to better distinguish cell types, reduces demands on cell number, and show that our method of finding variably-methylated regions finds biologically meaningful loci. Finally, we present a software tool, \texttt{scbs}, that implements this approach and offers further functionality requried to work with scBS data.

@@

Sequencing-based assays with single-cell resolution have offered new means to understand the differences between the cells making up a sample. Single-cell RNA sequencing (sc-RNA-seq) techniques have matured at great pace in recent years, with well developed analysis methodology, and methods to study epigenetics at single-cell resolution are rapidly catching up.

@@

\begin{figure*}
    \hspace{2.5cm}A\\[-7ex]
    \begin{center}
        \includegraphics[width=0.7\textwidth]{figures/Fig_residuals_A.png}\\
    \end{center}
    \hspace{2.5cm}B\\[-3ex]
    \begin{center}
        \includegraphics[width=0.7\textwidth]{figures/Fig_residuals_B.png}
    \end{center}
    \caption{\small \textbf{Effect of read position:} \textbf{(A)} Depicted is a genomic interval along a chromosome, for which DNA methylation is to be quantified. Two cells cover differing parts of the interval with one read each. If one simply counts for each cell which fraction of the covered CpG sites are methylated, one obtains very different values for the two cells. \textbf{(B)} By averaging each CpG site's methylation over all cells and subsequent smoothing, the thick black ``average methylation curve'' is obtained. To quantify the methylation of cell 1 from (A) relative to this average over all cells, we propose to use the cell's residuals to the smoothed curve (lengths of the vertical red lines) and take their average, counting residuals of methylated CpGs positive and residuals of unmethylated CpGs negative.}
    \label{fig:smoothres}
\end{figure*}

@@

\begin{figure*}
	\begin{center}
		\includegraphics[width=.8\textwidth]{figures/Fig_workflow.png}
	\end{center}
	\caption{\small \textbf{Overview of the functionalities implemented in the \texttt{scbs} package.}
		\texttt{scbs prepare} parses methylation files produced by common bisulfite sequencing mappers and stores their contents in a compressed format optimised for efficient access to genomic intervals.
		\texttt{scbs} also produces cell-wise summary statistics and quality plots (here: average methylation around the transcription start site) that are used to detect low-quality cells.
		These cells can be discarded with \texttt{scbs filter}.
		To obtain a methylation matrix, similar to the count matrices used in scRNA-seq, the user must first decide in which genomic intervals methylation should be quantified.
		The user may either provide genomic regions of \emph{a priori} interest, or they may choose to discover VMRs (variably-methylated regions) in the data with \texttt{scbs scan}.
		The resulting methylation matrix can then be used for downstream analysis such as cell clustering and dimensionality reduction.
	}
	\label{workflow}
\end{figure*}

>>>>

\usepackage[a4paper,hmargin=1.5cm,vmargin=2cm]{geometry}
\setlength{\columnsep}{0.7cm}
%\usepackage{microtype} % nice typesetting  % doesn't work for me :-( -S
\usepackage{amsmath}
\usepackage{graphicx}

\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\urlstyle{same} % no tt font for URLs
\usepackage{natbib}
\bibliographystyle{genome_research}
%\setcitestyle{aysep={}} 
\usepackage[dvipsnames]{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={blue!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!50!black}
}
\renewcommand{\topfraction}{0.9}
\renewcommand{\dbltopfraction}{0.9}
\renewcommand{\textfraction}{0.1}

@@

June 2022
\vspace{5.5ex}

\begin{minipage}{0.7\textwidth}
\small
\textit{Abstract} -- Single-cell-bisulfite sequencing (scBS) allows to asses DNA methylation with single-basepair and single-cell resolution. Analysing the big data sets thus obtained requires preprocessing to reduce data size, improve signal over noise, and provide interpretability. Currently, this is simply done by tiling the genome into large windows and averaging in these. We show that this coarse-graining dilutes signal, and suggest improved approaches to determine informative regions to quantify for and to quantitate methylation in these windows. We demonstrate how our preprocessing approach helps to better distinguish cell types, reduces demands on cell number, and show that our method of finding variably-methylated regions finds biologically meaningful loci. Finally, we present a software tool, \texttt{scbs}, that implements this approach and offers further functionality requried to work with scBS data.
\end{minipage}
\vspace{6.5ex}

}]

@@

Sequencing-based assays with single-cell resolution have offered new means to understand the differences between the cells making up a sample. Single-cell RNA sequencing (sc-RNA-seq) techniques have matured at great pace in recent years, with well developed analysis methodology, and methods to study epigenetics at single-cell resolution are rapidly catching up. 
\blfootnote{\hspace{-.5cm}\raggedright\textsuperscript{*} e-mail:  \texttt{l.kremer@dkfz-heidelberg.de} and\\ \hspace{1cm}\texttt{simon.anders@bioquant.uni-heidelberg.de}}

@@

\begin{figure*}[t]
    \hspace{2.5cm}A\\[-7ex]
    \begin{center}
        \includegraphics[width=0.7\textwidth]{figures/Fig_residuals_A.png}\\
    \end{center}
    \hspace{2.5cm}B\\[-3ex]
    \begin{center}
        \includegraphics[width=0.7\textwidth]{figures/Fig_residuals_B.png}
    \end{center}
    \caption{\small \textbf{Effect of read position:} \textbf{(A)} Depicted is a genomic interval along a chromosome, for which DNA methylation is to be quantified. Two cells cover differing parts of the interval with one read each. If one simply counts for each cell which fraction of the covered CpG sites are methylated, one obtains very different values for the two cells. \textbf{(B)} By averaging each CpG site's methylation over all cells and subsequent smoothing, the thick black ``average methylation curve'' is obtained. To quantify the methylation of cell 1 from (A) relative to this average over all cells, we propose to use the cell's residuals to the smoothed curve (lengths of the vertical red lines) and take their average, counting residuals of methylated CpGs positive and residuals of unmethylated CpGs negative.}
    \label{fig:smoothres}
\end{figure*}

@@

\begin{figure*}[t]
	\begin{center}
		\includegraphics[width=.8\textwidth]{figures/Fig_workflow.png}
	\end{center}
	\caption{\small \textbf{Overview of the functionalities implemented in the \texttt{scbs} package.}
		\texttt{scbs prepare} parses methylation files produced by common bisulfite sequencing mappers and stores their contents in a compressed format optimised for efficient access to genomic intervals.
		\texttt{scbs} also produces cell-wise summary statistics and quality plots (here: average methylation around the transcription start site) that are used to detect low-quality cells.
		These cells can be discarded with \texttt{scbs filter}.
		To obtain a methylation matrix, similar to the count matrices used in scRNA-seq, the user must first decide in which genomic intervals methylation should be quantified.
		The user may either provide genomic regions of \emph{a priori} interest, or they may choose to discover VMRs (variably-methylated regions) in the data with \texttt{scbs scan}.
		The resulting methylation matrix can then be used for downstream analysis such as cell clustering and dimensionality reduction.
	}
	\label{workflow}
\end{figure*}
